Approximation as Differentiation

### 2025-04-16 appendix/notes/grothendieck.tex
Clarify 'nudge' notation.
Minor correction.

\begin{definition}[Grothendieck construction for $F$]
Suppose a functor $F: \cat{C} \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{\cat{C}}F$ for
$F$ is the category where:
\begin{itemize}
\item objects $X$ are pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ paired with morphisms
$\faml{f}: F(f)(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
\end{itemize}
\end{definition}

\noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
$\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.

@@

\begin{enumerate}
\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(\cat{C})$ with $X' = \Fam(\cat{C})(X,Y)$ and $Y'_f =
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $(f, g): X \times Y \to Z$ to
$(f', g'): X \to [Y, Z]$ with:
\begin{itemize}
\item $f'(x) = (f \comp (x, -), \reindex{g}{(x,-)} \comp \biinj_{\faml{Y}}):
Y \to Z$ where $\biinj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\item $g'_x = \lambda_{\prod_{\idx{Y}}}(\reindex{g}{(x,-)} \comp \biinj_{\faml{X}_x}:
\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f})$ where $\biinj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\end{itemize}
\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(\cat{C})$ with
\begin{itemize}
\item $\idx{\eval_{X,Y}}(f, x) = \idx{f}(x)$
\item $\faml{\eval_{X,Y}}_{f, x} = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
\end{itemize}
\end{enumerate}

>>>>

\begin{definition}[Grothendieck construction for $F$]
Suppose a functor $F: \cat{C} \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{\cat{C}}F$ for
$F$ is the category which has:
\begin{itemize}
\item as objects $X$ all pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
\item as morphisms $f: X \to Y$ all pairs of a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ and a
morphism $\faml{f}: F(\idx{f})(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
\end{itemize}
\end{definition}

\noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
$\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to F(\idx{f})(\faml{Y})$ in $F(\idx{X})$.

@@

\begin{enumerate}
\item $[X, Y]$ to be the object $(X',Y')$ in $\Fam(\cat{C})$ with $X' = \Fam(\cat{C})(X,Y)$ and $Y'_f =
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}}$ for every $f: X \to Y$.
\item The family of isomorphisms $\lambda_{X,Y,Z}$ (natural in $X$) sending any $(f, g): X \times Y \to Z$ to
$(f', g'): X \to [Y, Z]$ with:
\begin{itemize}
\item $f'(x) = (f \comp (x, -), \reindex{g}{(x,-)} \comp \biinj_{\faml{Y}}):
Y \to Z$ where $\biinj_{\faml{Y}}: \faml{Y} \to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\item $g'_x = \lambda_{\prod_{\idx{Y}}}(\reindex{g}{(x,-)} \comp \biinj_{\faml{X}_x}:
\const_{\idx{Y}}(\faml{X}_x) \to \reindex{\faml{Z}}{f})$ where $\biinj_{\faml{X}_x}: \const_{\idx{Y}}(\faml{X}_x)
\to \const_{\idx{Y}}(\faml{X}_x) \biprod \faml{Y}$
\end{itemize}
Here $(x,-): \idx{Y} \to \idx{X} \times \idx{Y}$ denotes the function $\prodM{\const(x)}{\id_{\idx{Y}}}$ which
pairs any $y \in \idx{Y}$ with $x \in \idx{X}$.
\item The family of morphisms $\eval_{X,Y}: [X, Y] \times X \to Y$ in $\Fam(\cat{C})$ with
\begin{itemize}
\item $\idx{\eval_{X,Y}}(f, x) = \idx{f}(x)$
\item $\faml{\eval_{X,Y}}_{f, x} = \coprodM{{\eval_{\prod}}_x}{\faml{f}_x}:
\prod_{\idx{X}}\reindex{\faml{Y}}{\idx{f}} \biprod \faml{X}_x \to \faml{Y}_{\idx{f}(x)}$
\end{itemize}
\end{enumerate}
